---
title: figures_and_tables_for_text
output:
  html_document:
    df_print: paged
    number_sections: no
    self_contained: yes
    smooth_scroll: no
    toc: yes
    toc_depth: 4
    toc_float: no
    toc_position: top
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  warning = FALSE, message =  FALSE) 
```

```{r packeges}
library(tidyverse)
library(table1)
library(haven)
library(readxl)
library(stringr)
library(stringi)
library(ggalt)
library(ggrepel)
library(colourpicker)
```

```{r functions}
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g),correct = TRUE, simulate.p.value = TRUE, B = 10000)$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```

```{r data}
# Dissertations (Russian Book Chamber)
book <- read_csv("data/book.csv")
book <- book %>% mutate(type_diss = factor(type_diss, levels = c("PhD", "DS")))

# Raw data (before gender assignment) for Table 1
book_before_gender_assign <- read_csv("data/book_before_gender_assign.csv")
```

## Figures

#### Figure. 1 Share of female and male authors in 18 research fields

```{r, fig.align = "center", fig.dim = c(9, 4)}
plot_order <- book %>%
  count(level_1_en, gender) %>%
  group_by(level_1_en) %>%
  mutate(prop = prop.table(n)) %>%
  filter(gender == "female") %>%
  arrange(desc(prop))

lvls <- as.character(plot_order$level_1_en)

label_order <- book %>%
  count(level_1_en, type_diss, gender) %>%
  group_by(level_1_en,type_diss) %>%
  mutate(prop = prop.table(n)) 

book %>% 
  mutate(level_1_en = factor(level_1_en, levels = lvls)) %>%
  
  ggplot(aes(level_1_en, fill = gender)) +
  geom_bar(position = position_fill(reverse = TRUE), color = "white", alpha = 0.8) +
  facet_grid(~type_diss) + 
  geom_text(data = label_order, 
            aes(label = 100 * round(prop,digits = 2), y = prop), 
            position = position_stack(reverse = TRUE,vjust = 0.5),
            size = 3, alpha = 0.8)+
  scale_y_continuous(labels=scales::percent) +
  coord_flip() + 
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_fill_manual(values = c("#ffbe1a", "#b2b1b4")) +
  labs(x = "", y = "Figure. 1 Share of female and male authors in 18 research fields") +
  theme_bw() + 
  theme(legend.position = "top",
        legend.justification = 'left',
        legend.title = element_blank(),
        legend.text = element_text(size = 9),
        panel.grid = element_blank(),
        strip.background = element_rect(fill = NA),
        strip.text = element_text(colour = 'black', size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9)) +
  guides(fill=guide_legend(nrow = 1, byrow = TRUE)) +
  theme(plot.margin = unit(c(1, 0.2, 0.2, 0.2), "cm"))+
  theme(legend.position=c(0.001, 1.18))

ggsave(filename = "pic/fig1.png",
       units = "mm", dpi = 300,  width = 180,
       height = 100)
```

<br/>

#### Figure. 2 Gap between the share of female authors on PhD and DS levels

```{r, fig.align = "center", fig.dim = c(8, 4.2)}

df <- book %>%
  count(level_1_en, type_diss, gender) %>%
  group_by(level_1_en, type_diss) %>%
  mutate(prop = 100 * prop.table(n))  %>% 
  filter(gender == "female") 

plot_order <- df %>% select(-gender, -n) %>% 
  pivot_wider(names_from = type_diss, values_from = prop) %>% 
  mutate(gap = PhD - DS) %>% 
  arrange(desc(gap))

lvls <- as.character(plot_order$level_1_en)

ggplot() +  
  geom_vline(xintercept = 50, linetype = "dashed", alpha = 0.5, size = 0.4) +
  geom_rect(data = plot_order, 
            aes(xmin = 100, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "grey90") +
  geom_vline(xintercept = 100, size = 0.2) +
  geom_text(data = plot_order, 
            aes(x = 105, y = reorder(level_1_en, gap), label = round(gap, 1)), size = 3.2) +
  geom_dumbbell(data = plot_order, 
                aes(x = PhD, xend = DS, y = level_1_en, group = level_1_en),
                color = "grey80", size = 1.8,
                size_x = 3.2, size_xend = 3.2) + 
  geom_text(data = plot_order, 
            aes(x = PhD, label = format(round(PhD, digits = 0), nsmall = 0), y = level_1_en),
            color = "black", hjust = -0.8, size = 2.5) +
  geom_text(data = plot_order, 
            aes(x = DS, label = format(round(DS, digits = 0), nsmall = 0), y = level_1_en),
            color = "black", hjust = 1.8, size = 2.5) +
  geom_point(data = df, aes(prop, level_1_en, fill = type_diss), size = 4, colour = "black", pch = 21) +
  scale_fill_manual(values = c("#F95700FF", "#00A4CCFF"))+
  scale_x_continuous(limits = c(0, 105), labels = function(x) paste0(x, "%")) +
  labs(title = "Gap (pp)", 
       x = "Figure. 2 Gap between the share of female authors on PhD and DS levels") + 
  theme_bw()+ 
  theme(panel.grid.major = element_line(size=0.1, color = "grey80", linetype = "solid"),
        panel.grid.minor = element_line(size = 0.1, color = "grey80", linetype = "solid"),
        panel.spacing = unit(5, "mm"),
        plot.title = element_text(hjust = 1, size = 10),
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        legend.title = element_blank(),
        legend.text = element_text(size = 9)) +
  guides(fill = guide_legend(nrow = 1,byrow = TRUE)) +
  theme(plot.margin = unit(c(6.2, 5.5, 5.5, 5.5), "points")) +
  theme(legend.position = c(0.1, 1.07))

ggsave(filename = "pic/fig2.png",
       units = "mm", dpi = 300,  width = 160,
       height = 105)
```

<br/>

#### Figure. 3 Share of female authors of PhDs and DSs in 2005-2016

```{r, fig.align = "center", , fig.dim = c(9, 5)}
lvls <- book %>% filter(type_diss == "PhD") %>% 
  group_by(level_1_en, gender) %>% count() %>% 
  pivot_wider(names_from = gender, values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(tot = male + female) %>% 
  rowwise() %>% 
  mutate(ratio = 100 * female / tot) %>% arrange(desc(ratio)) %>% 
  mutate(level_1_en = as.factor(level_1_en)) %>% pull(level_1_en)

book %>%
  mutate(level_1_en = factor(level_1_en, levels = lvls)) %>% 
  group_by(year_letopis, type_diss, level_1_en, gender) %>% count() %>% 
  pivot_wider(names_from = gender, values_from = n) %>% 
  replace(is.na(.), 0) %>% 
  mutate(tot = male + female) %>% 
  rowwise() %>% 
  mutate(ratio = 100 * female/tot) %>% 
  
  ggplot(aes(year_letopis, ratio, color = type_diss, group = type_diss)) +
  geom_hline(yintercept = 50, color = "grey75", linetype = "dashed") +
  geom_line(alpha = 0.6) +
  geom_point(size = 0.8) +
  geom_text(aes(label = ifelse(type_diss == "PhD" & (year_letopis == "2005" | year_letopis =="2016"), 
                               paste0(round(ratio,  digits = 0), "%"), '')),
            size = 2.2, vjust = -1.5, color = "black") +
  scale_color_manual(values=c("#F95700FF", "#00A4CCFF")) +
  scale_y_continuous(limits = c(0, 110), breaks = c(0, 50, 100)) +
  scale_x_continuous(limits = c(2004, 2017)) +
  facet_wrap( ~ level_1_en, nrow = 3) +
  labs(y = "Share of women (%)", 
       x = "Figure. 3 Share of female authors of PhDs and DSs in 2005-2016") +
  theme_test()+
  theme(panel.grid.major = element_line(size = 0.1, color = "grey80", linetype = "solid"),
        panel.grid.minor = element_line(size = 0.1, color = "grey80", linetype = "solid"),
        legend.position = "top",
        legend.justification = 'left',
        legend.title = element_blank(),
        strip.background = element_rect(fill = NA),
        strip.text = element_text(colour = 'black', size = 8),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 8)) 

ggsave(filename = "pic/fig3.png",
       units = "mm", dpi = 300,  width = 200,
       height = 110)
```

<br/>

## Tables

#### Table 1. Research fields with counts of dissertations in 2005-2016. 

```{r}
tab_1 <- table1( ~ level_1_en,
        data = book_before_gender_assign)

write.table(tab_1, "tables/tab_1.csv", col.names = TRUE, row.names = FALSE, append = TRUE, sep = ',')
tab_1
```

<br/>

#### Table 2. Share of female and male authors of PhDs and DSs.

In the text for ISSI 2023 in Table 2 we have 227730 PhDs, here less by 5. It is because the `book` data here with the dropped observations with unknown organizations.

```{r}
tab_2 <- table1( ~ gender | type_diss,
        overall = FALSE,
        #extra.col = list(`P-value` = pvalue),
        data = book)

write.table(tab_2, "tables/tab_2.csv", col.names = TRUE, row.names = FALSE, append = TRUE, sep = ',')
tab_2
```

<br/>

